### 2020-06-29 paper.tex
added notes for paper

Notes for paper:
- list requirements and assumptions (e.g. Files are RO)
- make sure it's clearly explained why Sea over writing your own script to manage filesystems
    - Sea will manage disk and memory for you (abstraction of the script to be written)
    - Difficulties of intercepting libc calls. Describe and explain custom code that has been added
- Spark vs Burst Buffer vs Lustre caching vs Sea. Need to benchmark Sea vs burst buffer to measure
  applicability when Burst Buffer options are available
- Sea vs Neuroimaging workflow engines
- Splitting data into chunks to maximize storage space or defined constant with max file size
- Intercepting system calls vs libc calls.
- Making Sea flushing pipeline-aware through scheduler hooks
- Flushing: Pros and cons of single flusher/node vs multiple flushers/node (maximize network bandwidth usage)
- Distributed Sea? Sea will function in a distributed manner with Sub. Not for this paper